NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models

### 2018-03-27 paper/overview.tex

\section{NLIZE System}
In this section, we discuss the design and implementation of the NLIZE (pronounced as ``analyze'', see interface overview in Fig.~\ref{fig:teaser}) system, which address the analysis tasks (\textbf{T1-3}).
%
As discussed in the task analysis, the experts often analyze and obtain intuition about the model by studying how altering one part of the model affect other stages of the pipeline.
%
Such a process can be generalized as the perturbation-driven paradigm, which is used as guiding principal for designing the proposed tool.
%
As illustrated in Fig.~\ref{fig:modelPipeline}, we enables the automated or user-guided perturbation of the input sentence (e.g., replace words), the attention in the model (e.g., alter the alignment between sentences), and the prediction (e.g., adjust the prediction by making updates to the model).
%
In the following sections, we describe in detail the five major components of the proposed system, namely, sentence view  (Section~\ref{sec:sentence}), attention view (Section~\ref{sec:attention}), prediction view (Section~\ref{sec:prediction}), pipeline view (Section~\ref{sec:pipeline}), all pair summary view (Section~\ref{sec:allPairs}).

\begin{figure}[htbp]
\centering
 \includegraphics[width=1.0\linewidth]{pipeline}
 \caption{
Perturbation-driven exploration of the natural language inference model.
In the proposed tool, we enables the automated or user-guided perturbation of the input sentence (e.g., replace words), the attention in the model (e.g., alter the alignment between sentences), and the prediction (e.g., adjust the prediction by making updates to the model).
}
\label{fig:modelPipeline}
\end{figure}

>>>>

\section{NLIZE System}
In this section, we discuss the design and implementation of the NLIZE (pronounced as ``analyze'', see interface overview in Fig.~\ref{fig:teaser}) system, which address the visualization tasks (\textbf{T1-3}).
%
As discussed in the task analysis, the experts often analyze and obtain intuition about the model by studying how altering one part of the model affect other stages of the pipeline.
%
Such a process can be generalized as the perturbation-driven paradigm, which is used as guiding principal for designing the proposed tool.
%
As illustrated in Fig.~\ref{fig:modelPipeline}, we enables the automated or user-guided \emph{perturbation} (i.e., replace words) of the input sentence, the \emph{perturbation} of attention (i.e., alter the alignment between sentences)inside the model, as well as the \emph{perturbation} the prediction (e.g., adjust the prediction by making updates to the model).
%
In the following sections, we describe in detail the five major components of the proposed system, namely, sentence view  (Section~\ref{sec:sentence}), attention view (Section~\ref{sec:attention}), prediction view (Section~\ref{sec:prediction}), pipeline view (Section~\ref{sec:pipeline}), all pair summary view (Section~\ref{sec:allPairs}).

\begin{figure}[htbp]
\centering
 \includegraphics[width=1.0\linewidth]{pipeline}
 \caption{
 Perturbation-driven exploration of the natural language inference model.
 In the proposed tool, we enables the interrogation of the relationship between different components of the model via the perturbation-based analysis.
 %
 The user can perturb the of the input sentence (i.e., replace words with synonymous), perturb the attention (i.e., alter the soft alignment between sentences), and perturb the prediction (i.e., adjust the prediction by making updates to the model).
}
\label{fig:modelPipeline}
\end{figure}

### 2018-03-27 paper/prediction.tex

\subsection{Prediction View}
\label{sec:prediction}
For a given sentence pair, the model predicts a discrete probability distribution of the three labels (neutral, contradict, and entailment).
%
In the prediction view, as illustrated in Fig.~\ref{fig:predictionView}, a prediction probability is encoded as a point in the barycentric coordinate system of the triangle. In the triangle, the distinct color indicates the region corresponds to different labels. The prediction result for the original sentence pair is represented by the larger yellow circle, whereas smaller grey circles illustrate the perturbed sentence pairs. A density contour of the prediction is computed to emphasis the highly cluttered areas and distinguish the outliers.

To address \textbf{T3}, we need a way to communicate the reassignment of the prediction. As illustrated in (b), we integrate such operation in the prediction panel. When press and drag the circle (prediction), the user is presented with the three option (E, N, C) corresponds to the labels. When the user hovers on one of the options, a dotted line is shown to indicate the newly assigned label.

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{predictionView}
 \vspace{-2mm}
 \caption{
In the prediction view, the prediction is encoded as a point in the barycentric coordinate system of the triangle shown in (a).
%
The prediction result for the original sentence pair is represented by a larger yellow circle and the prediction of perturbed pairs are illustrated by smaller grey circles.
A density contour of the prediction is computed to emphasize the highly cluttered areas and distinguish the outliers.
As illustrated in (b), the predicted label does not match the ground truth, therefore, we apply a label reassignment operation, which triggers the model update.
% when press and drag the circle (prediction), the user is presented with the three option corresponds to the labels. The dotted line and arrow indicate the newly assigned label.
 }
\label{fig:predictionView}
\end{figure}

>>>>

\subsection{Prediction View}
\label{sec:prediction}
For a given sentence pair, the model predicts a discrete probability distribution of the three labels (neutral, contradict, and entailment).
%
In the prediction view, as illustrated in Fig.~\ref{fig:predictionView}(a), a prediction probability is encoded as a point in the barycentric coordinate system of the triangle. In the triangle, the distinctly colored background indicates the regions corresponds to different labels. The prediction result for the original sentence pair is represented by the larger yellow circle, whereas smaller grey circles illustrate the perturbed sentence pairs. A density contour of the prediction is computed to emphasize the highly cluttered areas and distinguish the outliers.

To perturb the prediction (\textbf{T3}), we need a way to communicate the reassignment of the predicted label. As illustrated in Fig.~\ref{fig:predictionView}(b), we integrate such operation in the prediction view. When press and drag the prediction (represented as a circle), the user is presented with the three options (E, N, C) correspond to the labels. When the user hovers on one of the options, a dotted line is shown to indicate the newly assigned label. The assignment is applied when the user releases the mouse while hovering on the label of choice.

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{predictionView}
 \vspace{-2mm}
 \caption{
In the prediction view, the prediction is encoded as a point in the barycentric coordinate system of the triangle shown in (a).
 %
% The prediction result for the original sentence pair is represented by a larger yellow circle and the prediction of perturbed pairs are illustrated by smaller grey circles.
A density contour of the prediction is computed to emphasize the highly cluttered areas and distinguish the outliers.
As illustrated in (b), the predicted label does not match the ground truth. Therefore, we apply a label reassignment operation, which triggers a model update.
% when press and drag the circle (prediction), the user is presented with the three option corresponds to the labels. The dotted line and arrow indicate the newly assigned label.
 }
\label{fig:predictionView}
\end{figure}
